Stateless Model Checking with Data-Race Preemption Points

### 2016-03-17 oopsla-16/evaluation.tex
Expand related work to address PLDI reviewer concerns.

We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for cvars,
\paradise~for semaphores,
and \rwldgr~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used {\tt without\_function} to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
We tested Pintoses with 3 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented,
	so each test could only be run on a subset of the 78 submissions; see ``Number tested'' in Table~\ref{tab:drbugs}.
}.
The source code of all 9 test cases is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on {\tt malloc}'s internal lock.
We ran all tests on 12-core 3.2 GHz Xeon machines. %with 12GB of RAM.
In total, our evaluation comprises 629 tests (pairs of a test program and a Pintos or P2),
at least 163 of which could expose bugs.

>>>>

We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for cvars,
\paradise~for semaphores,
and \rwldgr~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used the {\tt without\_function} command to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
We tested Pintoses with 3 programs from the class's provided test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented,
	so each test could only be run on a subset of the 78 submissions; see ``Number tested'' in Table~\ref{tab:drbugs}.
}.
The source code of all 9 test cases is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on {\tt malloc}'s internal lock.
We ran all tests on 12-core 3.2 GHz Xeon machines. %with 12GB of RAM.
In total, our evaluation comprises 629 unique tests (pairs of a test program and a Pintos or P2),
at least 163 of which could expose bugs.

### 2016-03-17 oopsla-16/overview.tex
Expand related work to address PLDI reviewer concerns.

\subsection{Data Race Detection}

@@

Stand-alone data-race analyses must avoid inundating the user with false alarms \cite{racerx}.
However, this work incorporates data-race analysis in an internal feedback loop,
using MC to automatically check each potential race
and reporting to the user only of directly observed failures.
%Our convergence theorem requires identifying all potential data-races to provide total verification (\sect{\ref{sec:totalverif}})
%% TODO CAMREADY: cite our soundness proofs tech report in this paragraph as well
Hence, we employ Limited HB in this paper},
accepting some overhead from false positives for the sake of total verification (\sect{\ref{sec:totalverif}}).

>>>>

\subsection{Data Race Detection}
\label{sec:overview-dr}

@@

Stand-alone data-race analyses must avoid inundating the user with false alarms \cite{racerx}.
However, this work incorporates data-race analysis in an internal feedback loop,
using MC to automatically check each potential race
and reporting to the user only directly observed failures.
%Our convergence theorem requires identifying all potential data-races to provide total verification (\sect{\ref{sec:totalverif}})
%% TODO CAMREADY: cite our soundness proofs tech report in this paragraph as well
Hence, we employ Limited HB in this paper},
accepting some overhead from false positives for the sake of total verification (\sect{\ref{sec:totalverif}}).

### 2016-03-17 oopsla-16/related.tex
Expand related work to address PLDI reviewer concerns.

{\bf Other techniques.} Various improvements or alternatives to DPOR have been developed, such as Dynamic Interface Reduction \cite{demeter}, Maximal Causality Reduction \cite{mcr},
%DPOR for relaxed memory models \cite{tsopso},
SAT-directed MC \cite{satcheck},
\revision{and R4 for event-driven applications \cite{r4}}.
These are all compatible with our technique.
Recent work \cite{tsopso} has extended DPOR for relaxed memory models \cite{memory-consistency-models},
which we do not yet account for in our proofs (\sect{\ref{sec:soundness}}).
Parrot \cite{parrot} combines MC with a partially-determinizing runtime for further reduction, but still, fewer than half the non-trivial state spaces in their evaluation could be completed.
%providing a strong argument for \quicksand.
Finally, Iterative Context Bounding (ICB) \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB is limited to a fixed set of PPs, and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan} combine the lockset and happens-before analyses into a hybrid technique, which we employ.
\revision{Several tools \cite{droidracer,cafa,unnamed-android} have recently emerged to target Android applications},
%DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%%event-driven
%Android applications,
using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} offers data-race techniques for kernel code.
IFRit \cite{ifrit}
improves performance using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
% No, IDGAF about pure happens before.
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-

>>>>

{\bf \revision{Reduction}~techniques.} Various improvements or alternatives to DPOR have been developed, such as Dynamic Interface Reduction \cite{demeter}, Maximal Causality Reduction \cite{mcr},
%DPOR for relaxed memory models \cite{tsopso},
SAT-directed MC \cite{satcheck},
\revision{and R4 for event-driven applications \cite{r4}}.
These are all compatible with our technique.
Recent work \cite{tsopso} has extended DPOR for relaxed memory models \cite{memory-consistency-models},
which we do not yet account for in our proofs (\sect{\ref{sec:soundness}}).
Parrot \cite{parrot} combines MC with a partially-determinizing runtime for further reduction, but still, fewer than half the non-trivial state spaces in their evaluation could be completed.
%providing a strong argument for \quicksand.

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT is optimized for tests with impossibly large state spaces,
while for more approachable tests,
its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
Finally, Iterative Context Bounding (ICB) \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification},
and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan}
%combine the lockset and happens-before analyses into a hybrid technique,
\revision{introduce Limited Happens-Before}, %(\sect{\ref{sec:overview-dr}})},
which we employ,
\revision{while \cite{predictive-dr} introduces the Causally-Precedes relation for a sound yet incomplete analysis (\sect{\ref{sec:overview-dr}})}.
\revision{Several tools \cite{droidracer,cafa,unnamed-android} have recently emerged to target Android applications},
%DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%%event-driven
%Android applications,
using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} \revision{finds data races in}~kernel code.
IFRit \cite{ifrit}
improves performance using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
% No, IDGAF about pure happens before.
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-
